---
title: "Iteration and List Columns"
output: github_document
---

```{r}
library(tidyverse)
library(rvest)

set.seed(1)
```


## Lists

```{r}
vec_numeric = 5:8
vec_char = c("My", "name", "is", "Jeff")
vec_logical = c(TRUE, TRUE, TRUE, FALSE)
```

```{r}
l = list(
  vec_numeric = 5:8,
  mat         = matrix(1:8, 2, 4),
  vec_logical = c(TRUE, FALSE),
  summary     = summary(rnorm(1000)))
l
```
Lists can be accessed using names or indices,
`$` extract things from list, generally avoid to.
```{r}
l$vec_numeric
```

```{r}
l[[1]]
```

```{r}
l[[1]][1:3]
```


## `for` loops

Make a different list
```{r}
list_norms = 
  list(
    a = rnorm(20, 3, 1),
    b = rnorm(20, 0, 5),
    c = rnorm(20, 10, .2),
    d = rnorm(20, -3, 1)
  )

is.list(list_norms)
```

Use function from last class again:
```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```

Apply function to the lists:
```{r}
mean_and_sd(list_norms[[1]])
```

```{r}
mean_and_sd(list_norms[[2]])
```

```{r}
mean_and_sd(list_norms[[3]])
```

```{r}
mean_and_sd(list_norms[[4]])
```

Use a loop to iterate:
```{r}
output = vector("list", length = 4) # define output as a type of list with length of 4


for (i in 1:4) {
  output[[i]] = mean_and_sd(list_norms[[i]])
}
output
```


##`map`

Use map to do the same thing
```{r}
output = map(list_norms, mean_and_sd)
      # map (the input list, function)

output
```

```{r}
output = vector("list", length = 4)

for (i in 1:4) {
  output[[i]] = median(list_norms[[i]])
}

output = map(list_norms, median)
output
```


## `map` variants

Use `map_dbl` because median outputs a *single* numeric value each time; the result is a vector instead of a list. 
Using the `.id` argument keeps the names of the elements(a/b/c/d) in the input list.
```{r}
output = map_dbl(list_norms, median, .id = "input")
output
```
If we tried to use `map_int` or `map_lgl`, we’d get an error because the output of median isn’t a integer or a logical. This is a good way to help catch mistakes when they arise.


Use the output-specific `map_dfr`; this will produce a single data frame, with `input` column of list.
```{r}
output = map_dfr(list_norms, mean_and_sd, .id = "input")
output
```
The `map_df` variants can be helpful when your map statement is part of a longer chain of piped commands.


The variant `map2` (and `map2_dbl`, etc) is helpful when your function has two arguments.
`
output = map2(input_1, input_2, \(x,y) func(arg_1 = x, arg_2 = y))
output
`


## Lists columns and operations

Try to put my list into a dataframe.
```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"),
    samp = list_norms
  )
```

Check if the list is working?
```{r}
listcol_df |> pull(name)
```

```{r}
listcol_df |> pull(samp)
```
Examine the first list entry using usual list index 
```{r}
pull(listcol_df, samp)[[1]]
```

Apply mean_and_sd to the first element of our list column.
```{r}
mean_and_sd(pull(listcol_df, samp)[[1]])
```

`pull(listcol_df, samp)` is a list. 
Iterate mean_and_sd function to each element using `map`
```{r}
map(pull(listcol_df, samp), mean_and_sd)
```
The `map` function returns a list; we could store the results as a new list column

Using `mutate` to define a new variable in a data frame/add a column, especially one that is a function of an existing variable
```{r}
listcol_df = 
  listcol_df |> 
  mutate(summary = map(samp, mean_and_sd))

listcol_df
```
`summary` column has list of mean and sd in `tibble` form.

Pull out `summary` column to check
```{r}
pull(listcol_df, summary)
```
`unnest` a particular column in the dataframe. 
```{r}
listcol_df |>
  select(-samp) |>
  unnest(summary)
```



## Revisiting NSDUH


```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

nsduh_table <- function(html, table_num) {
  
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent)) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  table
}

```

Try with a `for` loop:
```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

output = vector("list", 3)

for (i in c(1, 4, 5)) {
  output[[i]] = nsduh_table(nsduh_html, i)
}

nsduh_results = bind_rows(output)
output
```

Do this with `map`
```{r}
nsduh_results = 
  map(c(1, 4, 5), nsduh_table, html = nsduh_html) |> 
  bind_rows()
nsduh_results
```

Do this all in a dataframe
```{r}
nsduh_results= 
  tibble(
    name = c("marj", "cocaine", "heroine"), #name each table
    number = c(1, 4, 5)) |> 
  mutate(table = map(number, nsduh_table, html = nsduh_html)) |> 
  unnest(cols = "table")

nsduh_results
```

An alternative way to accomplish this is to use map’s syntax for “anonymous” (i.e. not named and saved) functions. Indicating that the only argument to care about is `num` and then passing that into `nsduh_html`.
```{r}
nsduh_results= 
  tibble(
    name = c("marj", "cocaine", "heroine"),
    number = c(1, 4, 5)) |> 
  mutate(
    table = map(number, \(num) nsduh_table(html = nsduh_html, num))) |> 
  unnest(cols = "table")

nsduh_results
```


## Operations on nested data


```{r}
library(p8105.datasets)
data("weather_df")
```

Weather data will be separated into three station-specific data frames, each of which is the data “observation” for the respective station.
```{r}
weather_nest = 
  nest(weather_df, data = date:tmin)

weather_nest
```

Different way of producing a `list` column. The result is a lot like `listcol_df`, in that the columns in weather_nest are a vector(for name) and a list(for data).
```{r}
weather_nest |> pull(name)
```

```{r}
weather_nest |> pull(data)
```

If you can `nest` data you should be able to `unnest` it as well, and you can (with the caveat that you’re unnesting a list column that contains a data frame).
```{r}
unnest(weather_nest, cols = data)
```


Let's do a simple regression:
```{r}
weather_df |>
  filter(name == "CentralPark_NY") |>
  lm(tmax ~ tmin, data = _)

weather_df |>
  filter(name == "Molokai_HI") |>
  lm(tmax ~ tmin, data = _)

weather_df |>
  filter(name == "Waterhole_WA") |>
  lm(tmax ~ tmin, data = _)
```

Instead, let's iterate differently:

Fit the simple linear regression relating `tmax` to `tmin` for each station-specific data frame.
```{r}
weather_lm = function(df) {
  lm(tmax ~ tmin, data = df)
}
```

Let’s make sure this works on a single data frame.
```{r}
weather_lm(pull(weather_nest, data)[[1]])
```

Also, can do this with `map`:
```{r}
map(pull(weather_nest, data), weather_lm)
```

Again, to avoid the creation of a dedicated function using `map`’s syntax for “anonymous” (i.e. not named and saved) functions. This is helpful for really short operations – here, rather than writing a separate (named) function that takes the argument df and returns a regression of tmax on tmin.

I can use `\(df)` to indicate that same function. If you have a one-line function, this is a good option; otherwise creating a named function is probably better.
```{r}
map(pull(weather_nest, data), \(df) lm(tmax ~ tmin, data = df))
```

Let’s use mutate to fit this model, and to store the result in the same dataframe.
```{r}
weather_nest = 
  weather_nest |> 
  mutate(models = map(data, weather_lm))

weather_nest
```


## Using iteration for data import

Create a dataframe that includes the list of all files in that directory and the complete path to each file. 
Then, map over paths and import data using the read_csv function. 
Finally, unnest the result of map.
```{r}
full_df = 
  tibble(
    files = list.files("data/exp_data/"),
    path = str_c("data/exp_data/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest()
```

Tides the data using string manipulations on the file, converting from wide to long, and selecting relevant variables.
```{r}
tidy_df = 
  full_df %>% 
  mutate(
    files = str_replace(files, ".csv", ""),
    group = str_sub(files, 1, 3)) %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "outcome",
    names_prefix = "week_") %>% 
  mutate(week = as.numeric(week)) %>% 
  select(group, subj = files, week, outcome)
```

Finally, the code chunk below creates a plot showing individual data, faceted by group.
```{r}
tidy_df %>% 
  ggplot(aes(x = week, y = outcome, group = subj, color = group)) + 
  geom_point() + 
  geom_path() + 
  facet_grid(~group)
```







