---
title: "Simulation"
output: github_document
---

Load key package and source necessary files.
```{r}
library(tidyverse)
set.seed(1)

source("sim_mean_sd.R")

```

We can "simulate" by running our function.
```{r}
sim_mean_sd(n_subj = 400)
```


## Simulation: Mean and SD for one N

```{r}
sim_mean_sd = function(n_subj, mu = 2, sigma = 3) {
  
  sim_data = tibble(
    x = rnorm(n = n_subj, mean = mu, sd = sigma),
  )
  
  sim_data |> 
    summarize(
      mu_hat = mean(x),
      sigma_hat = sd(x)
    )
}
```

Can I verify the Central Limit Theorem?

First try with a 'for' loop
```{r}
output = vector("list", 100)

for (i in 1:100) {
  output[[i]] = sim_mean_sd(30)
}

sim_results = bind_rows(output)
```

```{r}
sim_results
```

```{r}
sim_results |> 
  ggplot(aes(x = mu_hat)) + 
  geom_density()
```

Try to repeat with a 'map' statement. 

```{r}
sim_results_df = 
  tibble(iter = 1:100) |>
  mutate(results = map(iter, sim_mean_sd, n_subj = 30)) |>
  unnest(results)

sim_results_df
```

## Data frame with multiple iterations for multiple sample sizes
```{r}
sim_results_df = 
  expand_grid(
    sample_size = 30,
    iter = 1:100
  ) |> 
  mutate(
    estimate_df = map(sample_size, sim_mean_sd)
  ) |> 
  unnest(estimate_df)

sim_results_df
```

```{r}
sim_results_df = 
  expand_grid(
    sample_size = c(30,60,90,120),
    iter = 1:100
  ) |> 
  mutate(
    estimate_df = map(sample_size, sim_mean_sd)
  ) |> 
  unnest(estimate_df)

sim_results_df
```

Let's look at this:
```{r}
sim_results_df |>
  mutate(
    sample_size = str_c("n = ", sample_size), 
    sample_size = fct_inorder(sample_size)
  ) |>
  ggplot(aes(x = sample_size, y = mu_hat)) +
  geom_violin()
```

Let's try to summarize

```{r}
sim_results_df |> 
  pivot_longer(
    mu_hat:sigma_hat,
    names_to = "parameter", 
    values_to = "estimate") |> 
  group_by(sample_size) |> 
  summarize(
    emp_mean = mean(estimate),
    emp_sd = sd(estimate)) |> 
  knitr::kable(digits = 3)
 

```


## Simulation: Mean for several 𝑛


```{r}
sim_results_df = 
  expand_grid(
    sample_size = c(30, 60, 120, 240),
    iter = 1:1000
  ) |> 
  mutate(
    estimate_df = map(sample_size, sim_mean_sd)
  ) |> 
  unnest(estimate_df)
```


```{r}
sim_results_df |> 
  mutate(
    sample_size = str_c("n = ", sample_size),
    sample_size = fct_inorder(sample_size)) |> 
  ggplot(aes(x = sample_size, y = mu_hat, fill = sample_size)) + 
  geom_violin()
```

```{r}
sim_results_df |> 
  pivot_longer(
    mu_hat:sigma_hat,
    names_to = "parameter", 
    values_to = "estimate") |> 
  group_by(parameter, sample_size) |> 
  summarize(
    emp_mean = mean(estimate),
    emp_var = var(estimate)) |> 
  knitr::kable(digits = 3)
```


## Simulation: Simple Linear Regression for one 𝑛

```{r}
sim_df = 
  tibble(
    x = rnorm(30, mean = 1, sd = 1),
    y = 2 + 3 * x + rnorm(30, 0, 1)
  )

sim_df |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point()

slr_fit = lm(y ~ x, data = sim_df)

coef(slr_fit)
```


turn this into a function

```{r}
sim_regression = function(n_subj, beta_0 = 2, beta_1 = 3) {
  
  sim_df = 
    tibble(
      x = rnorm(n_subj, mean = 1, sd = 1),
      y = beta_0 + beta_1 * x + rnorm(n_subj, 0, 1)
    )

  slr_fit = lm(y ~ x, data = sim_df)

  tibble(
    beta0_hat = coef(slr_fit)[1],
    beta1_hat = coef(slr_fit)[2]
  )
}
```

```{r}
sim_regression(n_subj = 30)
```


```{r}
output = vector("list", length = 500)

for (i in 1:500) {
  
  output[[i]] = sim_regression(n_subj = 30)
  
}

output |> 
  bind_rows()


```

```{r}
slr_sim_results_df = 
  expand_grid(
    sample_size = 30, 
    iter = 1:500
  ) |> 
  mutate(
    results = map(sample_size, sim_regression)
  ) |> 
  unnest(results)

slr_sim_results_df |> 
  ggplot(aes(x = beta0_hat, y = beta1_hat)) + 
  geom_point()
```

## One more example!


```{r}
birthdays = sample(1:365, 5, replace = TRUE)

repeated_bday = length(unique(birthdays)) < 5

repeated_bday
```

put this in a function

```{r}
bday_sim = function(n_room) {
  
  birthdays = sample(1:365, n_room, replace = TRUE)

  repeated_bday = length(unique(birthdays)) < n_room

  repeated_bday
  
}

bday_sim(20)
```


```{r}
bday_sim_results = 
  expand_grid(
    bdays = 5:50, 
    iter = 1:2500
  ) |> 
  mutate(
    result = map_lgl(bdays, bday_sim)
  ) |> 
  group_by(
    bdays
  ) |> 
  summarize(
    prob_repeat = mean(result)
  )
```

plot this

```{r}
bday_sim_results |> 
  ggplot(aes(x = bdays, y = prob_repeat)) + 
  geom_point() + 
  geom_line()
```



## Varying two simulation parameters


```{r}
sim_results_df = 
  expand_grid(
    sample_size = c(30, 60, 120, 240),
    true_sd = c(6, 3),
    iter = 1:1000
  ) |> 
  mutate(
    estimate_df = 
      map2(sample_size, true_sd, \(n, sd) sim_mean_sd(n = n, sigma = sd))
  ) |> 
  unnest(estimate_df)
```

```{r}
sim_results_df |> 
  mutate(
    true_sd = str_c("True SD: ", true_sd),
    true_sd = fct_inorder(true_sd),
    sample_size = str_c("n = ", sample_size),
    sample_size = fct_inorder(sample_size)) |> 
  ggplot(aes(x = sample_size, y = mu_hat, fill = sample_size)) + 
  geom_violin() + 
  facet_grid(. ~ true_sd)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```























